\section{OrientDB}
The orientDB  is a distributed Multi-Model NoSQL database having an additional
Graph Database Engine \cite{hid-sp18-412-orientDB_by_CallidusCloud}. In order
to support the efficient control of the data and
discover the true potential of the data orientDB provides the support for the
multi-model in contrast to the primitive single-model database. The novel
multi-model approach of the orientDB has enabled the users to work
simultaneously with the Graph, Document, key-value, geo-Spatial and
reactive models. The multi-model database drives to reduce the
complications with respect to the operations and maintain the consistency
in the data \cite{hid-sp18-412-orientDB_multimodel}.

The orientDB is the NoSQL graph database which utilizes the graph data model. 
The graph is modeled with the vertices and the edges. The vertices can be a 
person, place, object or a chunk of the data. The relation between the two 
nodes can be represented by an edge \cite{hid-sp18-412-orientDB_graph}. 
The orientDB has the custom JDBC driver that helps the user to connect 
to the remote servers. This enables the users to interact with the 
database by utilizing the standard set of interactions in the java. 
The orientDB is compatible with the Tinkerpop API. This API provides 
the effective functionalities to abstract the graph, vertex, edge 
and the other properties \cite{hid-sp18-412-orientDB_graph}.

The database provides the support for the searching
the data of any type with the several indexing mechanisms based on the
B-tree and the extensible hashing. The orientDB also supports the three 
types of the schema models, namely schema-less, schema-full and the 
schema-mixed modes. Each of the record in the database has the 
surrogate key that points to the record in the ArrayList. In the DB-Engine 
graph database standings orientDB stands 
in the 3rd position \cite{hid-sp18-412-orientDB_wiki}.






